\subsubsection{Simulated Annealing}

In the third difficulty, 
we use Simulated Annealing to plan the route for the monster AI.
Although Simulated Annealing can not guarantee finding the shortest path,
it can still be used to obtain a near-optimal path, 
which is sufficient for our Monster AI. In this project, the purpose of SA is
not only to chase the human efficiently, but also to make the monster's route
less rigid than a purely deterministic shortest-path algorithm. Therefore, the
SA planner is designed to keep enough pursuit pressure while still allowing
small variations between similar candidate paths.

Initially, the initial state is built by greedy search
with toroidal Manhattan distance as its heuristic function.
However, after some tests, we found that in some cases,
greedy search might fail to reach the target,
causing the entire simulated annealing process to fail.
Therefore, we used Depth-First Search(DFS) to generate the initial state.
The DFS prefers moves that are closer to the human using toroidal Manhattan distance,
but it also shuffles candidates so the path is not completely deterministic.
This means that the initial path is usually directed toward the human, while
still preserving some diversity in the order of exploration. Several tests
proved that this metric is effective enough to reach the target state in most
valid maps, and the time complexity is also acceptable because the DFS gives
priority to cells with smaller Manhattan distance.

The Simulated Anealing generates neibour solution by cutting the current path at a random point 
and trying to rebuild the remaining path from that cut point to the human. The
prefix before the cut point is kept unchanged, and the suffix is rebuilt by the
same DFS-based method. If the rebuilt suffix reaches the human, the candidate
path becomes a new complete route. If the search cannot reconnect to the human
within the limit, the truncated path can still be evaluated. This design is
useful for gameplay because the monster only executes the first two steps of
the planned path in each round, so a candidate route can still provide a useful
short-term movement decision even if its far endpoint is not perfect.
The cost of each state is calculated by the following function:
\[
E(P)=Cost(P)-J(P)+Repeat(P)+\Delta(P).
\]
The first term is the chase cost:
\[
Cost(P)=|P|+3D_M(p_{\mathrm{move}},h)+0.5D_M(p_{\mathrm{end}},h).
\]
Here, \(|P|\) keeps the planned path short, \(p_{\mathrm{move}}\) is the
position reached after the monster's two actual movement steps, and
\(p_{\mathrm{end}}\) is the last point of the planned path. \(D_M(\cdot,h)\) is
the toroidal Manhattan distance to the human. \(D_M(p_{\mathrm{move}},h)\) is
added to encourage the monster to move towards the human, while
\(D_M(p_{\mathrm{end}},h)\) is added to penalize routes that fail to reach the
target. Since only the first two steps of the planned route are used in each
round, the second term is given a larger weight than the third term. This makes
the score focus on whether the move made in the current round increases the
monster's pressure on the human, rather than requiring every candidate path to
be a fully optimal route to the target.

The second term rewards junction control:
\[
J(P)=\min\left(6,\sum_i j_i\right),
\qquad
j_i=
\begin{cases}
2.0, & \deg(p_i)\ge 3\text{ and }i<6,\\
0.5, & \deg(p_i)\ge 3\text{ and }i\ge 6,\\
0, & \text{otherwise,}
\end{cases}
\]
where \(\deg(p_i)\) is the number of walkable neighboring cells. Since
\(J(P)\) is also subtracted, the monster is encouraged to pass through
junctions, which gives it more possible future routes and makes its movement
harder for the player to predict. The first few cells receive a larger reward
because they are the cells that can affect the monster's immediate behaviour in
the current turn. The cap of \(6\) prevents a long path from gaining too much
score advantage simply by passing through many junctions.

The third term uses the previous SA move:
\[
Repeat(P)=10I(p_{\mathrm{move}}=p_{\mathrm{prev,start}}).
\]
This term penalizes immediately moving back to the previous starting cell. It
adds a small amount of cross-turn variation without punishing the monster for
continuing in the same direction when that direction is still a good chase
choice. In other words, the monster is allowed to keep pursuing directly, but
it is discouraged from visibly oscillating between two cells.

Finally, the fourth term adds small route variation:
\[
\Delta(P)=2\cdot \frac{1}{k}\sum_{i=0}^{k-1} \delta(p_i),
\qquad
k=\min(|P|,6),
\]
where \(\delta(p_i)\in[-1,1]\) is a random cell cost generated once at the start of
the current SA run. This term is applied only to the first six cells because the
monster only moves two steps after planning, so early route variation matters
more than changes near the end of the path. The weight of this term is small,
so it mainly changes the choice between paths with similar chase quality. It
does not dominate the chase cost, which prevents the monster from wandering
randomly.

Overall, the objective still minimizes path length and strongly prefers paths
that move toward the human. The random, junction, and previous-move terms are
deliberately smaller, so they mainly act as tie breakers between similar chase
paths rather than making the monster wander randomly.

The SA parameters were tuned with two goals: the monster decision should be fast
enough for real-time play, and the optimized path should still be shorter than
the DFS initial path in most generated maps. The original setting used
\(INNER\_LOOP=200\), an initial temperature of \(1000\), a minimum temperature
of \(0.001\), and a cooling rate of \(0.995\). This produced about \(551600\)
path mutations in one SA call. Although this setting could improve the path
quality, the average wall-clock time on five generated maps was about
\(3807.27\) ms, which was too slow for interactive gameplay.

Therefore, we reduced the search budget step by step. The main parameters were
\(INNER\_LOOP\), the initial temperature, the minimum temperature, the cooling
rate, and \(DFS\_EXTEND\_SEARCH\_LIMIT\). Reducing \(INNER\_LOOP\) and using a
faster cooling rate directly lowered the number of candidate paths tested.
However, making the search too small also reduced the benefit of SA. For
example, a very small configuration with about \(165\) mutations improved only
one out of five maps, while a configuration with about \(450\) mutations still
improved three out of five maps but was less reliable on difficult layouts.

The final configuration uses \(INNER\_LOOP=40\), initial temperature \(100\),
minimum temperature \(0.1\), cooling rate \(0.97\), and
\(DFS\_EXTEND\_SEARCH\_LIMIT=80\). This configuration performs about \(9080\)
path mutations per SA call. In a test on five newly generated maps, all five
runs finished within one second, with an average wall-clock time of
\(280.00\) ms. The average DFS path length was \(65.4\), while the average SA
path length was reduced to \(40.2\). This shows that the final parameters keep
the algorithm responsive while preserving enough search quality to improve the
initial DFS route.
